---
output: html_notebook
---
## 4.4 For-loops

The formula for the sum of the series $1+2+...+n$ is $n(n+1)/2$. What if we weren't sure that was the right function? How could we check? Using what we learned about functions we can create one that computes the $S_n$:

```{r}
compute_s_n <- function(n){
  x <- 1:n
  
  sum(x)
}
```

How can we compute &S_n& for various values on &n&, say $n = 1, ..., 25$? Do we write 25 lines of code calling `compute_s_n`? No, that is what for-loops are for in programming. In this case, we are performing exactly the same task over and over, and the only thing that is changing is the value of $n$. For-loops let us define the range that our variable takes (in our example $n = 1, ..., 10$), then change the value and evaluate expression as you $loop$.


Perhaps the simplest example of a for-loop is this useless piece of code:

```{r}
for (i in 1:5){
  print(i)
}
```

Here is the for-loop we would write for our $S_n$ example:

```{r}
m <- 25
s_n <- vector(length = m) # create an empty vector
for(n in 1:m){
  s_n[n] <- compute_s_n(n)
}
```

In each iteration $n=1, n=2,$ etc..., we compute $S_n$ and store it in the $n$th entry of `s_n`.

Now we can create a plot to search for a pattern:

```{r}
n <- 1:m
plot(n, s_n)
```

If you noticed that it appears to be a quadratic, you are on the right track because the formula is $n(n+1)/2$, which we can confirm with a table:

```{r}
head(data.frame(s_n = s_n, formula = n*(n+1)/2))
```

We can also overlay the two results by using the function `lines` to draw a line over the previously plotted points:

```{r}
plot(n, s_n)
lines(n, n*(n+1)/2)
```

## 4.5 Vectorization and functionals

Although for-loops are an important concept to understand, in R we rarely use them. As you learn more R, you will realize _vectorization_ is preferred over for-loops since it results in shorter and clearer code. We already saw examples in the Vector Arithmetic Section. A _vectorized_ function is a function that will apply the same operation on each of the vectors.

```{r}
x <- 1:10
sqrt(x)
```
```{r}
y <- 1:10
x*y
```

To make this calculation, there is no need for for-loops. However, not all functions work this way. For instance, the function we just wrote, `compute_s_n`, does not work element-wise since it is expecting a scalar. This piece of code does not run the function on each entry of `n`:

```{r}
n <- 1:25
compute_s_n(n)
```

_Functionals_ are functions that help us apply the same function to each entry in a vector, matrix, data frame or list. Here we cover the functional that operates on numeric, logical and character vectors: `sapply`.

The function `sapply` permits us to perform element-wise operations on any function. Here is how it works:

```{r}
x <- 1:10
sapply(x, sqrt)
```
> However, sqrt itself seems to be applicable to element-wise operation.

```{r}
x <- 1:10
sqrt(x)
```
> They show same result. What happened?

Other functions are `apply`, `lapply`, `tapply`, `mapply`, `vapply`, and `replicate`. We mostly use `sapply`, `apply` and `replicate` in this book, but we recomment familiarizing yourselves with the others as they can be very useful.

## 4.6 Exercises

1.\ What will this conditional expression return?

```{r}
x <- c(1,2,-3,4)

if(all(x>0)){
  print("All Positives")
} else{
  print("Not all positives")
}
```

2.\ Which of the following expressions is always `FALSE` when at least one entry of a logical vector `x` is TRUE?

A. `all(x)`

B. `any(x)`

C. `any(!x)`

D. `all(!x)`

Answer: D

3.\ The function `nchar` tells you how many characters long a character vector is.

Write a line of code that assigns to the object `new_names` the state abbreviation when the state name is longer than 8 characters.

Answer:
우선 `state.name`중 `nchar`가 8 보다 큰 주의 index를 추출한다.

```{r}
which(nchar(state.name)>8)
```

이를 이용해 `state.abb`에서 indexing해낸다.
```{r}
new_names <- vector(length = m)
new_names <- state.abb[which(nchar(state.name)>8)]
new_names
```

4.\ Create a function `sum_n` that for any given value, say `n`, computes the sum of the integers from 1 to n (inclusive). Use the function to determine the sum of integers from 1 to 5,000.

```{r}
sum_n <- function(n){
  sum(1:n)
}
```

```{r}
sum_n(5000)
```
여기에 plot으로 한 번 확인해보자.

```{r}
l <- 5000
s_hwagin <- vector(length = l)
for (n in 1:l){
  s_hwagin[n] <- sum_n(n)
}
```
```{r}
plot(s_hwagin)
b <- 1:5000
a <- function(n){
  n*(n+2)/2
}
lines(a(b))
```



5.\ Create a function `altman_plot` that takes two arguments, `x` and `y`, and plots the difference against the sum.

```{r}
altman_plot <- function(x, y){
  if (x >= y){
    plot(x-y, x+y )
  }
  else{
    plot(y-x, x+y)
  }
}
altman_plot(1, 2)
```
6.\ After running the code below, what is the value of x?

```{r}
x <- 3
my_func <- function(y){
  x <- 5
  y+5
}

x
```

7.\ Write a function `compute_s_n` that for any given $n$ computes the sum $S_n = 1^2 + 2^2+3^2+...n^2$. Report the value of the sum when $n=10$.

```{r}
compute_s_n <- function(n) {
  value <- vector("numeric", length = n) # Create a vector that stores $n^2$s as numeric elements.
 
  #for loop to assign n^2 to each index of [n]
  for(i in 1:n){
    value[i] <- i*i
  }
  return(sum(value))
}

```
```{r}
compute_s_n(10)
```

8.\ Define an empty numerical vector `s_n` of size 25 using `s_n <- vector("numeric", 25)` and store in the results of $S_1, S_2, ... S_{25}$using a for loop.

```{r}
s_n <- vector("numeric", 25)
for (i in 1:25) {
  s_n[i] <- compute_s_n(i)
}

s_n
```

9.\ Repeat exercise 8, but this time use `sapply`.

`sapply`는 함수를 interable object의 element에 대해 element-wise하게 각각 적용할 수 있게 해준다. 벡터, 리스트, 표현식, 데이터프레임을 받아 각각의 요소에 함수를 적용해주고 벡터나 행렬로 반환해준다.

```{r}

x <- 1:25
sapply(x, compute_s_n)
```

10.\ Repeat exercise 8, but this time use `map_dbl`.

`map_dbl`은 `purrr` library의 function이다. `sapply`와 비슷하다. for-loop의 대안으로 
```{r}
library(purrr)
x <- 1:25
map_dbl(x, compute_s_n)
```

11.\ Plot `S_n` versus `n`. Use points defined by $n=1,...,25$.

위에서 `1:25`를 `x`로 assign해주었으므로 n 대신 x를 쓴다.
```{r}
plot(x, s_n)
```

12.\ Confirm that the formula for this sum is $S_n = n(n+1)(2n+1)/6$

```{r}
expected_value <- function(n) {
  n*(n+1)*(2*n+1)/6
}
plot(x, s_n)
lines(x, expected_value(x)) #Lines함수의 경우 인자는 (x, y,...) 순이다. 이 때 y에는 x에대한 함수가 들어가야 한다.
```

> 그냥 함수를 쓰지 않고 `lines`안의 argument에 `x*(x+1)*(2*x+1)/6`을 직접 써줘도 동작하지 않을까? I don't see why not.

```{r}
plot(x, s_n)
lines(x, x*(x+1)*(2*x+1)/6)
```

정상적으로 동작하는 것을 확인했다! :)