## 3.7 Vectors

In R, the most basic objects available to store data are _vectors_. As we have seen, complex datasets can usually be broken down into components that are vectors. For example, in a data frame, each column is a vector. Here we learn more about this important class.


### 3.7.1 Creating vectors

We can create vectors using the function `c`, which stands for _concatenate_. We use `c` to concatenate entries in the following way:

```{r}
codes <- c(380, 124, 818)
codes
```

We can also create character vectors. We use the quotes to denote that the entries are characters rather than variable names.

```{r}
country <- c("italy", "canada", "egypt")
```

In R you can also use single quotes:

```{r}
country <- c('italy', 'canada', 'egypt')
```

But be careful not to confuse the single quote ' with the _back_ quote `.

By now you should know that if you type:

```{r, eval=FALSE}
country <- c(italy, canada, egypt)
```

you receive an error because the variables `italy`, `canada` and `egypt` are not defined. If we do not use the quotes, R looks for variables with those names and returns an error.


### 3.7.2 Names

Sometimes it is useful to name the entries of a vector. For example, when defining a vector of country codes, we can use the names to connect the two:

```{r}
codes <- c(italy = 380, canada = 124, egypt = 818)
codes
```

The object `codes` continues to be a numeric vector:

```{r}
class(codes)
```

but with names:

```{r}
names(codes)
```

If the use of strings without quotes looks confusing, know that you can use the quotes as well:

```{r}
codes <- c("italy" = 380, "canada" = 124, "egypt" = 818)
codes
```

There is no difference between this function call and the previous one. This is one of the many ways in which R is quirky compared to other languages.

We can also assign names using the `names` functions:

```{r}
codes <- c(380, 124, 818)
country <- c("italy", "canada", "egypt")
names(codes) <- country
codes
```


### 3.7.3 Sequences

Another useful function for creating vectors generates sequences:

```{r}
seq(1, 10)
```

The first argument defines the start, and the second defines the end which is included. The default is to go up in increments of 1, but a third arguments lets us tell it how much to jump by:

```{r}
seq(100, 1, -1)
```

If we want consecutive integers, we can use the following shorthand:

```{r}
1:10
```

When we use these functions, R produces integers, not numerics, because they are typicaly used to index something:

```{r}
class(seq(1, 10))
```
```{r}
class(1:10)
```

However, if we create a sequence including non-integers, the class changes:
```{r}
class(seq(1, 10, 0.5))
```


### 3.7.4 Subsetting

We use square brackets to access specific elements of a vector. For the vector `codes` we defined above, we can access the second element using:

```{r}
codes[2]
```

You can get more than one entry by using a multi-entry vector as an index:
```{r}
codes[c(1,3)]
```
The sequences defined above are particularly useful if we want to access, say, the first two elements:

```{r}
codes[1:2]
```
If the elements have names, we can also access the entries using these names. Below are two examples.

```{r}
codes["canada"]
```
```{r}
codes[c("egypt","italy")]
```

## 3.8 Coercion

In general, _coercion_ is an attempt by R to be flexible with data types. When an entry does not match the expected, some of the prebuilt R functions try to guess what was meant before throwing an error. This can also lead to confusion. Failing to understand _coercion_ can drive programmers crazy when attempting to code in R since it behaves quite differently from most other languages in this regard. Let's learn about it with some examples.

We said that vectors must be all of the same type. So if we try to combine, say, numbers and characters, you might expect an error:

```{r}
x <- c(1, "canada", 3)
```

But we don't get one, not even a warning! What happened? Look at `x` and its class:

```{r}
x
```
```{r}
class(x)
```

R _coerced_ the data into characters. It guessed that because you put a character string in the vector, you meant the 1 and 3 to actually be character strings "`1`" and "`3`". The fact that not even a warning is issued is an example how coercion can cause many unnoticed errors in R.

R also offers functions to change from one type to another. For example, you can turn numbers into characters with:

```{r}
x <- 1:5
y <- as.character(x)
y
class(y)
```

You can turn it back with `as.numeric`:
```{r}
as.numeric(y)
```

This function is actually quite useful since datasets that include numbers as character strings are common.

### 3.8.1 Not availables (NA)

When a function tries to coerce one type to another and encounters an impossible case, it usually gives us a warning and turns the entry into a special value called an `NA` for "not available". For example:

```{r}
x <- c("1", "b", "3")
as.numeric(x)
```

R does not have any guesses for what number you want when you type `b`, so it does not try.

As a data scientist you will encounter the `NA`s often as they are generally used for missing data, a common problem in real-world datasets.

*another examples*
R does not have any guess as to what "two" means as a number, unless we set `two` as a variable and assign a numeric value of `2`.
```{r}
x <- c("1", "two", "3")
as.numeric(x)
```

## 3.9 Exercises

1.\ Use the function `c` to create a vector with the average high temperatures in January for Beijing, Lagos, Paris, Rio de Janeiro, San Juan and Toronto, which are 35, 88, 42, 84, 81, and 30 degrees Fahrenheit. Call the object `temp`.

A.
```{r}
temp <- c(35, 88, 42, 84, 81, 30)
cities <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
names(temp) <- cities
temp
```

2.\ Now create a vector with the city names and call the object `city`.

```{r, eval=FALSE}
# Did just that above. :)
```

3.\ Use the `names` function and the objects defined in the previous exercises to associate the temperature with its corresponding city.

```{r, eval=FALSE}
# Also did just that :)
```

4.\ Use the `[` and `:` operators to access the temperature of the first three cities on the list.

```{r}
temp[1:3]
```

5.\ Use the `[` operator to access the temperature of Paris and San Juan.
```{r, eval=FALSE}
temp["Paris", "San Juan"]
#Notice that placing a comma in between `names` in strings does not subset the two elements.
```

Instead, the names should also be a character vector.

```{r}
temp[c("Paris", "San Juan")]
```

6.\ Use the `:` operator to create a sequence of numbers $12, 13, 14, ..., 73$.

```{r}
12:73
```
We can also use `seq()` function in order to create such a sequence. Its last argument is "by", and its default is 1 if you leave it blank.
```{r}
seq(12, 73)
```

7.\ Create a vector containing all the positive odd numbers smaller than 100.

```{r}
seq(1, 99, 2)
```

8.\ Create a vector of numbers that starts at 6, does not pass 55, and adds numbers in increments of 4/7: 6, 6+4/7, 6+8/7, etc.. How many numbers does the list have? Hint: use `seq` and `length`.

```{r}
length(seq(6, 55, 4/7))
```

9.\ What is the class of the following object `a <- seq(1, 10, 0.5)`?

It must be numeric. `seq(integer.a, integer.b)` will result in integer because `seq()` function is widely used in counting. However, setting `by` argument as `0.5` results in numeric values and therefore changes the class of elements into numeric. Therefore it is a numeric vector.
Let's test it out.

```{r}
sequence <- seq(1, 10, 0.5)
class(sequence)
```

10.\ What is the class of the following object `a <- seq(1, 10)`?

Unlike the above case, it is an integer vector.

```{r}
sequence1 <- seq(1, 10)
class(sequence1)
```

11.\ The class of `class(a<-1)` is numeric, not integer. R defaults to numeric and to force an integer, you need to add the letter `L`. Confirm the class that the class of `1L` is integer.

```{r}
class(a<-1)
class(1L)

# One can also transform the class of a by using `as.integer()` function.
class(as.integer(a))
```

12.\ Define the following vector:

```{r}
x <- c("1", "3", "5")
```
and coerce it to get integers.

```{r}
as.integer(x)
```
```{r}
class(as.integer(x))
```

## 3.10 Sorting

Now that we have mastered some basic R knowledge, let's try to gain some insights into the safety of different states in the context of gun murders.

### 3.10.1 `sort`

Say we want to rank the states from least to most gun murders. The function `sort` sorts a vector in increasing order. We can therefore see the largest number of gun murders by typing:

```{r}
library(dslabs)
data(murders)
sort(murders$total)
```

However, this does not give us information about which states have which murder totals. For example, we don't know which state had 1257.

### 3.10.2 `order`

The funtion `order` is closer to what we want. It takes a vector as input and returns the vector of indexes that sorts the input vector. This may sound confusing so let's look at a simple example. We can create a vector and sort it:

```{r}
x <- c(31, 4, 15, 92, 65)
sort(x)
```

Rather than sort the input vector, the function `order` returns the index that sorts input vector:

```{r}
index <- order(x)
index
x[index]
```

This is the same output as that returned by `sort(x)`. If we look at this index, we see why it works:

```{r}
x
```

```{r}
order(x)
```

The second entry of `x` is the smallest, so `order(x)` starts with `2`. The next smallest is the third entry, so the second entry is `3` and so on.

How does this help us order the states by murders? First, remember that the entries of vectors you access with $ follow the same order as the rows in the table. For example, these two vectors containing state names and abbreviations respectively are matched by their order:

```{r}
murders$state[1:10]
```
This means we can order the state names by their total murders. We first obtain the indes that orders the vectors according to murder totals and then index the state names vector:

```{r}
ind <- order(murders$total)
murders$abb[ind]
```
According to the above, California had the most murders.

### 3.10.3 `max ` and `which.max`

If we are only interested in the entry with the largest value, we can use `max` for the value:

```{r}
max(murders$total)
```

and `which.max` for the index of the largest value:

```{r}
i_max <- which.max(murders$total)
murders$state[i_max]
```

For the minimum, we can use `min` and `which.min` in the same way.

Does this mean California the most dangerous state? In an upcoming section, we argue that we should be considering rates instead of totals. Before doing that, we introduce one last order-related function: `rank`.

### 3.10.4 `rank`

Although not as frequently used as `order` and `sort`, the function `rank` is also related to order and can be useful. For any given vector it returns a vector with the rank of the first entry, second entry, etc., of the input vector. Here is a simple example:

```{r}
x <- c(31, 4, 15, 92, 65)
rank(x)
```

To summarize, let's look at the results of the three functions we have introduced:

```{r}
data.frame('original' = x, 'sort' = sort(x), 'order' = order(x), 'rank' = rank(x))
```

### 3.10.5 Beware of recycling

Another common source of unnoticed errors in R is the use of _recyling_. We saw that vectors are added elementwise. So if the vectors don't match in length, it is natural to assume that we should get an error. But we don't. Notice what happens:

```{r}
x <- c(1, 2, 3)
y <- c(10, 20, 30, 40, 50, 60, 70)
x+y
```

We do get a warning but no error. For the output, R has recycled the numbers in `x`. Notice the last digit of numbers in the output.

# 3.11 Exercise

For these exercises we will use the US murders dataset. Make sure you load it prior to starting.

```{r}
library(dslabs)
data("murders")
```

1.\ Use the `$` operator to access the population size data and store it as the object `pop`. Then use the `sort` function to redefine `pop` so that it is sorted. Finally, use the `[` operator to report the smallest population size.

```{r}
sort(pop <- murders$population)[1]
```

2.\ Now instead of the smallest population size, find the index of the entry with the smallest population size. Hint: use `order` instead of `sort`.

```{r}
order(pop)[1]
```

3.\ We can actually perform the same operation as in the previous exercise using the function `which.min`. Write one line of code that does this.

```{r}
which.min(pop)
```

4. Now we know how small the smallest state is and we know which row represents it. Which state is it? Define a variable `states` to be the state names from the `murders` data frame. Report the name of the state with the smallest population.

```{r}
states <- murders$state
pop <- murders$population
states[which.min(pop)]
```

5.\ You can create a data frame using the `data.frame` function. Here is a quick example:
```{r}
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
city_temps <- data.frame('name' = city, 'temperature' = temp)
city_temps
```

Use the `rank` function to determine the population rank of each state from smallest population size to biggest. Save these ranks in an object called `ranks`, then create a data frame with the state name and its rank. Call the data frame `my_df`.

```{r}
ranks <- rank(murders$population)
my_df <- data.frame('name' = murders$state, 'rank' = ranks)
my_df
```

What if I reordered this for the sake of simplified look? 

```{r}
ordered_states <- murders$state[c(rev(order(rank(murders$population))))]
# It looks too messy. Let's do some revision on this afterwards.

my_df1 <- data.frame("state" = ordered_states, "population rank" = 1:51)
my_df1

```

6.\ Repeat the previous exercise, but this time order `my_df2` so that the states are orderd from least populous to most populous. Hint: create an object `ind` that stores the indexes needed to order the population values. Then use the bracket operator `[` to re-order each column in the data frame.

```{r}
ind <- order(murders$population)
states <- murders$state[ind]
pop <- murders$population[ind]
ranks <- rank(pop)
my_df2 <- data.frame("state"= states, "population" = pop, "rank" = ranks)

my_df2

#Introducing `ind` makes everything a bit more coherent and concise, unlike my own messy solution right above this one.
```

7.\ The `na_example` vector represents a series of counts. You can quickly examine the object using:

```{r}
data("na_example")
str(na_example)
```
However, when we compute the average with the function `mean`, we obtain `NA`:
```{r}
mean(na_example)
```
The `is.na` function returns a logical vector that tells us which entries are `NA`. Assign this logical vector to an object called `ind` and determine how many `NA`s does `na_example` have.
```{r}
ind <- is.na(na_example)
table(ind)["TRUE"]
```
8.\ Now compute the average again, but only for the entries that are not `NA`. Hint: remember the `!` operator.
```{r}
# na_example is an integer vector with NAs in it.

na_example 
```

