---
output: html_notebook
---

## 3.7 Vectors

In R, the most basic objects available to store data are _vectors_. As we have seen, complex datasets can usually be broken down into components that are vectors. For example, in a data frame, each column is a vector. Here we learn more about this important class.


### 3.7.1 Creating vectors

We can create vectors using the function `c`, which stands for _concatenate_. We use `c` to concatenate entries in the following way:

```{r}
codes <- c(380, 124, 818)
codes
```

We can also create character vectors. We use the quotes to denote that the entries are characters rather than variable names.

```{r}
country <- c("italy", "canada", "egypt")
```

In R you can also use single quotes:

```{r}
country <- c('italy', 'canada', 'egypt')
```

But be careful not to confuse the single quote ' with the _back_ quote `.

By now you should know that if you type:

```{r}
country <- c(italy, canada, egypt)
```

you receive an error because the variables `italy`, `canada` and `egypt` are not defined. If we do not use the quotes, R looks for variables with those names and returns an error.


### 3.7.2 Names

Sometimes it is useful to name the entries of a vector. For example, when defining a vector of country codes, we can use the names to connect the two:

```{r}
codes <- c(italy = 380, canada = 124, egypt = 818)
codes
```

The object `codes` continues to be a numeric vector:

```{r}
class(codes)
```

but with names:

```{r}
names(codes)
```

If the use of strings without quotes looks confusing, know that you can use the quotes as well:

```{r}
codes <- c("italy" = 380, "canada" = 124, "egypt" = 818)
codes
```

There is no difference between this function call and the previous one. This is one of the many ways in which R is quirky compared to other languages.

We can also assign names using the `names` functions:

```{r}
codes <- c(380, 124, 818)
country <- c("italy", "canada", "egypt")
names(codes) <- country
codes
```


### 3.7.3 Sequences

Another useful function for creating vectors generates sequences:

```{r}
seq(1, 10)
```

The first argument defines the start, and the second defines the end which is included. The default is to go up in increments of 1, but a third arguments lets us tell it how much to jump by:

```{r}
seq(100, 1, -1)
```

If we want consecutive integers, we can use the following shorthand:

```{r}
1:10
```

When we use these functions, R produces integers, not numerics, because they are typicaly used to index something:

```{r}
class(seq(1, 10))
```
```{r}
class(1:10)
```

However, if we create a sequence including non-integers, the class changes:
```{r}
class(seq(1, 10, 0.5))
```


### 3.7.4 Subsetting

We use square brackets to access specific elements of a vector. For the vector `codes` we defined above, we can access the second element using:

```{r}
codes[2]
```

You can get more than one entry by using a multi-entry vector as an index:
```{r}
codes[c(1,3)]
```
The sequences defined above are particularly useful if we want to access, say, the first two elements:

