---
output: html_notebook
---

# Chapter 5 The tidyverse 
_함수형프로그래밍이란 무엇인가_ 참조[^1] </br>
_Cheat sheets for purrr package_ 참조[^2]


[^1]: <https://statkclee.github.io/parallel-r/ds-fp-purrr.html>
[^2]: <https://github.com/rstudio/cheatsheets/blob/master/purrr.pdf>

Up to now we have been manipulating vectors by reordering and subsetting them through indexing. However, once we start more advanced analyses, the preferred unit for data storage is not the vector but the data frame. In this chapter we learn to work directly with data frames, which greatly facilitate the organization of information.. We will be using data frames for the majority of this book. We will focus on a specific data format referred to as _tidy_ and on specific collection of packages that are particularly helpful for working with tidy data referred to as the _tidyverse_.

We can load all the tidyverse packages at once by installing and loading the tidyverse package:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
```

We will learn how to implement the tidyverse approach throughout the book, but before delving into the details, in this chapter we introduce some of the most widely used tidyverse functionality starting with the **dplyr** package for manipulating data frames and the **purrr** package for working with functions. Note that the tidyverse also includes a graphing package, **ggplot2**, which we introduce later in Chpater 8 in the Data Visualiation part of the book; the ***readr* package discussed in Chapter 6; and many others. In this chapter, we first introduce the concept of _tidy_ data and then demonstrate how we use the tidyverse to work with data frames in this format.

## Tidy data

We say that a data table is in _tidy_ format if each row represents one observation and columns represent the different variables available for each of these observations. The `murders` dataset is an example of a tidy data frame.

```{r}
library(dslabs)
data(murders)
head(murders)
```

Each row represent a state with each of the five columns providing a different variable related to these states: name, abbreviation, region, population, and total murders.

To see how the same information can be provided in different formats, consider the following example:

```{r}
library(dslabs)
data("gapminder")
tidy_data <- gapminder %>% 
  filter(country %in% c("South Korea", "Germany") & !is.na(fertility)) %>%
  select(country, year, fertility)
head(tidy_data, 6)
```

This tidy dataset provides fertility rates for two countries across the years. This is a tidy dataset because each row presents one observation with the three variables being county, year, and fertility rate. However, this dataset originally came in another format and was reshaped for the `dslabs` package. Originally, the data was in the following format:

```{r}
#system.file의 extdata 함수를 이용하여 package에 포함된 데이터의 경로를 모두 알아내 path에 할당한다.
path <- system.file("extdata", package="dslabs")

#위의 path에서 file.path를 통해 "fertility-two-countries-example.csv"파일의 경로를 얻어 filename에 할당한다.
filename <- file.path(path, "fertility-two-countries-example.csv")

#read_csv는 csv파일을 데이터프레임으로 읽어들인다. 이를 wide_data에 할당한다.
wide_data <- read_csv(filename)

#wide_data에서 column 중 country와  1960:1962 의 데이터만 데이터프레임으로 출력한다.
select(wide_data, country, `1960`:`1962`) %>% as.data.frame
```

> R console의 `Parsed with column specification:`란은 각 column이 어떤 종류의 데이터를 포함하는지를 알려준다. </br>
> select(): Extract one or multiple columns as a data table. It can be also used to remove columns from the data frame. ex.) select(.data, ...) </br>
> select_if(): Select columns based on a particular condition. One can use this function to, for example, select columns if they are numeric. </br>
> pull(): Extract column values as a vector. The column of interest can be specified either by name or by index. </br>

The same information is provided, but there are two important differences in the format: 1) each row includes several observations and 2) one of the variables, year, is stored in the header. For the tidyverse packages to be optimally used, data need to be reshaped into `tidy` format, which you will learn to do in the Data Wrangling part of the book. Until then, we will use example datasets that are already in tidy format.

Although not immediately obvious, as you go through the book you will start to appreciate the advantages of working in a framework in which functions use tidy formats for both inputs and outputs. You will see how this permits the data analyst to focus on more important aspects of the analysis rather than the format of the data.

## 5.2 Exercises

1.\ Examine the built-in dataset `co2`. Which of the following is true:

A. `co2` is tidy data: it has one year for each row.

B. `co2` is not tidy: we need at least one column with a character vector.

C. `co2` is not tidy: it is a matrix instead of a data frame.

D. `co2` is not tidy: to be tidy we would have to wrangle it to have three columns (year, month and value), then each co2 observation would have a row.

Answer: D. `co2`라는 time series object는 year에 해당하는 column이 있도록 wrangle해줘야 tidy data가 된다.

```{r}
data(co2)
sprintf("Data type of co2 is %s", class(co2))
co2
```
`ts` 는 time-series object를 뜻한다. Data frame도, matrix도 아니다. Data frame으로 만들기 위해서는
`ts_reshape(ts.obj, type = "wide", frequency = NULL)`

를 이용할 수 있다.

```{r}
matrix_example <- matrix(1:12, nrow=3, ncol=3)
class(matrix_example)
```

2.\ Examine the built-in dataset `ChickWeight`. Which of the following is true:

A. ChickWeight is not tidy: each chick has more than one row.

B. ChickWeight is tidy: each observation (a weight) is represented by one row. The chick from which this measurement came from is one the variables.

C. ChickWeight is not a tidy: we are missing the year column.

D. ChickWeight is tidy: it is stored in a data frame.

```{r}
data("ChickWeight")
head(ChickWeight, 20)
```
Answer: B. ChickWeight is tidy: each observation (a weight) is represented by one row. The chick from which this measurement came from is one the variables.
Hint: 모든 변수를 column으로 표현해서 데이터를 쭉 적어내려가면 tidy 형태이다.

3.\ Examine the built-in dataset `BOD`. Which of the following is true:

A. BOD is not tidy: it only has six rows.

B. BOD is not tidy: the first column is just an index.

C. BOD is tidy: each row is an observation with two values (time and demand)

D. BOD is tidy: all small datasets are tidy by definition.

```{r}
data(BOD)
head(BOD)
```
Answer: C. 모든 row가 두 variable에 대한 observation이기 때문에 이 dataset은 tidy하다.

4.\ Which of the following built-in datasets is tidy (you can pick more than one):

A. `BJsales`

B. `EuStockMarkets`

C. `DNase`

D. `Formaldehyde`

E. `Orange`

F. `UCBAdmissions`

```{r, echo=FALSE}
data(BJsales)
print("BJsales")
BJsales
```
It's a time-series object with no variables as columns. Therefore `BJsales` is **NOT** a tidy data.

```{r}
data(EuStockMarkets)
str(EuStockMarkets)
```
Time-series의 시간(1991~)에 해당하는 variable column이 있어야 tidy format이다. `EuStockmarket`은 tidy data가 아니다.

```{r}
data(DNase)
head(DNase, 20)
```
DNase run 회차도 variable로서 column에 있으며, 각 row에는 run/conc/density 세가지 variable에 대하여 observed value를 기록해놓았다. 따라서 tidy data이다.

```{r}
data(Formaldehyde)
Formaldehyde
```

Variable에 해당하는 column과 observation이 기록된 row로 볼 때, `Formaldehyde`는 그냥 tidy data이다.

```{r}
data(Orange)
head(Orange, 20)
```
`Orange`는 tidy data이다.

```{r}
data("UCBAdmissions")
UCBAdmissions
```

tidy data가 아니다.

Answer: C, D, E

## 5.3 Manipulating data frames

The **dplyr** package from the tidyverse introduces functions that perform some of the most common operations when working with data frames and uses names for these functions that are relatively easy to remember. For instance, to change the data table by adding a new column, we use `mutate`. To filter the data table to a subset of rows, we use `filter`. Finally, to subset the data by selecting specific columns, we use `select`.

정리: 

- `mutate`: `mutate(.data, "name"= {values})`형태로 새로운 column (variable)을 추가한다.
- `filter`: `filter(.data, {conditional})`의 형태로 조건에 맞는 row만 subset해낸다.
- `select`: `select(.data, {variable names(column)})`의 형태로 원하는 cloumn만 추출할 수 있다.

### 5.3.1 Adding a column with `mutate`

We want all the necessary information for our analysis to be included in the data table. So the first task is to add the murder rates to our murders data frame. The function `mutate` takes the data frame as a first argument and the name and values of the variable as a second argument using the convention `name = values`. So, to add murder rates, we use:

```{r}
library(dslabs)
data(murders)
murders <- mutate(murders, rate = total / population * 100000)
```
Notice that here we used total and population inside the function, which are objects that are **not** defined in our workspace. But why don’t we get an error?

This is one of **dplyr**’s main features. Functions in this package, such as `mutate`, know to look for variables in the data frame provided in the first argument. In the call to mutate above, `total` will have the values in `murders$total` (likewise, `muders$population` for `population` ). This approach makes the code much more readable.

We can see that the new column is added:

```{r}
head(murders)
```

Although we have overwritten the original `murders` object, this does not change the object that loaded with `data(murders)`. If we load the `murders` data again, the original will overwrite our mutated version.

## 5.3.2 Subsetting with `filter`

Now suppose that we want to filter the data table to only show the entries for which the murder rate is lower than 0.71. To do this we use the `filter` function, which takes the data table as the first argument and then the conditional statement as the second. Like `mutate`, we can use the unquoted variable names from `murders` inside the function and it will know we mean the columns and not objects in the workspace.

```{r}
filter(murders, rate <= 0.71)
```

## 5.3.3 Selecting columns with `select`

Although our data table only has six columns, some data tables include hundreds. If we want to view just a few, we can use the **dplyr** `select` function. In the code below we select three columns, assign this to a new object and then filter the new object:

```{r}
new_table <- select(murders, state, region, rate)
filter(new_table, rate <= 0.71)
```

## 5.4  Exercises

1.\ Load the **dplyr** package and the murders dataset.
```{r}
library(dplyr)
library(dslabs)
data(murders)
```
Add columns using the dplyr function `mutate`. This function is aware of the column names and inside the function you can call them unquoted:
```{r}
murders <- mutate(murders, population_in_millions = population / 10^6)
```
We can write `population` rather than `murders$population`. The function `mutate` knows we are grabbing columns from `murders`.

Use the function `mutate` to add a murders column named `rate` with the per 100,000 murder rate as in the example code above. Make sure you redefine murders as done in the example code above ( murders <- [your code]) so we can keep using this variable.

```{r}
murders <- mutate(murders, rate = total / population * 10^5)
#오른쪽 mutate 항에서 column 생성. <-로 murders에 해당 column 붙여주기.
```
2.\ If `rank(x)` gives you the ranks of `x` from lowest to highest, `rank(-x)` gives you the ranks from highest to lowest. Use the function `mutate` to add a column `rank` containing the rank, from highest to lowest murder rate. Make sure you redefine `murders` so we can keep using this variable.

```{r}
murders <- mutate(murders, rank = rank(-rate))
```

3.\ With **dplyr**, we can use select to show only certain columns. For example, with this code we would only show the states and population sizes:

```{r, eval=FALSE}
select(murders, state, population) %>% head()
```
Use `select` to show the state names and abbreviations in `murders`. Do not redefine `murders`, just show the results.

```{r}
select(murders, state, abb) %>% head() #Piping해서 결과값을 head()에 집어넣는다.
```

4.\ The **dplyr** function filter is used to choose specific rows of the data frame to keep. Unlike select which is for columns, `filter` is for rows. For example, you can show just the New York row like this:

```{r, eval= FALSE}
filter(murders, state == "New York")
```

You can use other logical vectors to filter rows.

Use `filter` to show the top 5 states with the highest murder rates. After we add murder rate and rank, do not change the murders dataset, just show the result. Remember that you can filter based on the `rank` column.

```{r}
top_five_unordered <- filter(murders, rank(-rate) <= 5)
top_five <- top_five_unordered[order(top_five_unordered$rank), ]
top_five
```
Q. 왜 order로 정렬 후 5개 variable밖에 나타나지 않는것인가?
</br>
A. indexing에서 첫번째 argument가 row이며, `,`로 row와 column subset하는 것 잊지 말자!


```{r}
t5 <- filter(murders, rank(-rate) <= 5)
t5_ordered <- t5[c(order(t5$rank)), ]
t5_ordered
```

5.\ We can remove rows using the `!=` operator. For example, to remove Florida, we would do this:
```{r, eval=FALSE}
no_florida <- filter(murders, state != "Florida")
```

Create a new data frame called no_south that removes states from the South region. How many states are in this category? You can use the function `nrow` for this.

```{r}
library(dslabs)
library(dplyr)
data(murders)

no_south <- filter(murders, region != "South")
nrow(no_south)
```

6.\ We can also use %in% to filter with **dplyr**. You can therefore see the data from New York and Texas like this:

```{r}
filter(murders, state %in% c("New York", "Texas"))
```

Create a new data frame called `murders_nw` with only the states from the Northeast and the West. How many states are in this category?

```{r}
murders_nw <- filter(murders, region %in% c("Northeast", "West"))
nrow(murders_nw)
```
Answer: 22 states.

7.\ Suppose you want to live in the Northeast or West **and** want the murder rate to be less than 1. We want to see the data for the states satisfying these options. Note that you can use logical operators with `filter`. Here is an example in which we filter to keep only small states in the Northeast region.

```{r, eval = FALSE}
filter(murders, population < 5000000 & region == "Northeast")
```

Make sure `murders` has been defined with `rate` and `rank` and still has all states. Create a table called `my_states` that contains rows for states satisfying both the conditions: it is in the Northeast or West and the murder rate is less than 1. Use `select` to show only the state name, the rate and the rank.

```{r}
murders <- mutate(murders, rate = total / population * 10^5)
murders <- mutate(murders, rank = rank(-rate))

my_states <- filter(murders, (region == "Northeast" | region == "West") & rate < 1)

select(my_states, state, rate, rank)
```

Now that I learned the pipe operator `%>%` in the next section, I should try it out on this example.
```{r}
murders %>% filter(region == "Northeast" | region == "West") %>%
filter(rate < 1) %>% select(state, rate, rank)
```


## 5.5 The pipe: `%>%`

Tip: `%>%` can be called with keyboard shortcut: shift + ctrl + M.

With **dplyr** we can perform a series of operations, for example `select` and then `filter`, by sending the results of one function to another using what is called the _pipe operator_. `%>%`. Some details are included below.

We wrote code above to show three variables (state, region, rate) for states that have murder rates below 0.71. To do this, we defined the intermediate object `new_table`. In **dplyr** we can write code that looks more like a description of what we want to do without intermediate objects:

$$\textrm{original data} \rightarrow \textrm{select} \rightarrow \textrm{filter}$$

For such an operation, we can use the pipe `%>%`. The code looks like this:

```{r}
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
```

This line of code is equivalent to the  two lines of code above. What is going on here?

In general, the pipe _sends_ the result of the left side of the pipe to be the first argument of the function on the right side of the pipe. Here is a very simple example:

```{r}
16 %>% sqrt()
```

We can continue to pipe values along:

```{r}
16 %>% sqrt() %>%  log2()
```

The above statement is equivalent to `log2(sqrt(16))`.

Remember that the pipe sends values to the first argument, so we can define other arguments as if the first argument is already defined:

```{r}
16 %>% sqrt() %>% log(base = 2)
```

Therefore, when using the pipe with data frames and **dplyr**, we no longer need to specify the required first argument since the **dplyr** functions we have described all take the data as the first argument. In the code we wrote:

```{r, eval = FALSE}
murders %>%  select(state, region, rate) %>% filter(rate <= 0.71)
```
`murders` is the first argument of the `select` function, and the new data frame (formerly `new_table`) is the first argument of the `filter` function.

Note that the pipe works well with function where the first argument is the input data. Functions in **tidyverse** packages like **dplyr** have this format and can be used easily with the pipe.

## 5.6 Exercises

1.\ The pipe `%>%` can be used to perform operations sequentially without having to define intermediate objects. Start by redefining murder to include rate and rank.

```{r, warning = FALSE}
library(dplyr)
murders <- mutate(murders, rate = total / population * 10^5, rank = rank(-rate))
```

In the solution to the previous exercise, we did the following:

```{r, eval = FALSE}
my_states <- filter(murders, region %in% c("Northeast", "West") & rate < 1)

select(my_states, state, rate, rank)
```

The pipe `%>%` permits us to perform both operations sequentially without having to define an intermediate variable `my_states`. We therefore could have mutated and selected in the same line like this:

```{r, eval = FALSE}
mutate(murders, rate = total / population * 10^5, rank = rank(-rate)) %>% 
  select(state, rate, rank)
```

Notice that `select` no longer has a data frame as the first argument. The first argument is assumed to be the result of the operation conducted right before the `%>%`.

Repeat the previous exercise, but now instead of creating a new object, show the result and only include the state, rate, and rank columns. Use a pipe `%>%` to do this in just one line.

```{r}
mutate(murders, rate = total / population * 10^5, rank = rank(-rate)) %>% 
  filter((region == "Northeast" | region == "West") & rate < 1) %>% 
  select(state, rate, rank)
```

2.\ Reset `murders` to the original table by using `data(murders)`. Use a pipe to create a new data frame called `my_states` that considers only states in the Northeast or West which have a murder rate lower than 1, and contains only the state, rate, and rank columns. The pipe should also have four components separated by three `%>%`. The code should look something like this:

```
my_states <- murders %>% 
  mutate SOMETHING %>% 
  filter SOMETHING %>% 
  select SOMETHING
```

```{r}
library(dslabs)
data(murders)

my_states <- murders %>% 
  mutate(rate = total / population * 10^5, rank = rank(-rate)) %>%
  filter((region == "Northeast" | region == "West") & rate < 1) %>% 
  select(state, rate, rank)
my_states
```

## 5.7 Summarizing data

An important part of exploratory data analysis is summarizing data. The average and standard deviation are two examples of widely used summary statistics. More informative summaries can often be achieved by first splitting data into groups. In this section, we cover two new **dplyr** verbs that make these computations easier: `summarize` and `group_by`. We learn to acceses resulting values using the `pull` function.

### 5.7.1 `summarize`

The `summarize` function in **dplyr** provides a way to compute summary statistics with intuitive and readable code. We start with a simple example based on heights. The `heights` dataset includes heights and sex reported by students in an in-class survey.

```{r, message = FALSE}
library(dplyr)
library(dslabs)
data(heights)
```

The following code computes the average and standard deviation for females:

```{r}
s <- heights %>%
  filter(sex == "Female") %>% 
  summarize(average = mean(height), standard_deviation = sd(height))
s
```

Below is exactly the same thing, except the pipe is inside the `summarize` function as an argument, which defeats the purpose of using `%>%`... 

But whatever. :)
```{r}
summarize(heights %>%  filter(sex == "Female"), average = mean(height), std = sd(height))
```

This takes our original data table as input, filters it to keep only females, and then produces a new summarized table with just the average and the standard deviation of heights. We get to choose the names of the columns of the resulting table. For example, above we decided to use `average` and `standard_deviation`, but we could have used other names just the same.

Because the resulting table stored in `s` is a data frame, we can access the components with the accessor `$`:

```{r}
s$average
```
```{r}
s$standard_deviation
```

As with most other **dplyr** functions, `summarize` is aware of the variable names and we can use them directly. So when inside the call to the `summarize` function we write `mean(height)`, the function is accessing the column with the name "height" and then computing the average of the resulting numeric vector. We can compute any other summary that operates on vectors and returns a single value. For example, we can add the median, min and max heights like this:

```{r}
heights %>% 
  filter(sex == "Female") %>% 
  summarize(median = median(height), minimum = min(height), maximum = max(height))
```

We can obtain these three values with just one line using the `quantile` function: for example, `quantile(x, c(0,0.5,1))` returns the min (0th percentile), median (50th percentile), and max (100th percentile) of the vector `x`. However, if we attempt to use a function like this that returns two or more values inside `summarize`:

```{r}
heights %>% 
  filter(sex == "Female") %>% 
  summarize(range = quantile(height, c(0, 0.5, 1)))
```
we will receive an error: `Error: Column `range` must be length 1 (a summary value), not 3`. With the function `summarize`, we can only call functions that return a single value. In Section 5.12, we will learn how to deal with functions that return more than one value.

For another example of how we can use the `summarize` function, let's compute the average murder rate for the United States. Remember our data table includes total murders and population size for each state and we have already use **dplyr** to add a murder rate column:

```{r}
data(murders)
murders <- murders %>% mutate(rate = total/population * 100000)
```

Remember that the US murder rate is **not** the average of the state murder rates:

```{r}
summarize(murders, mean(rate))
```

This is because in the computation above the small states are given the same weight as the large ones.

The US murder rate is the total number of murders in the US divided by the total US population. So the correct computation is:

```{r}
us_murder_rate <- murders %>% 
  summarize(rate = sum(total)/sum(population) * 100000)
us_murder_rate
```

### 5.7.2 `pull`

The `us_murder_rate` object defined above represents just one number. Yet we are storing it in a data frame:
```{r}
class(us_murder_rate)
```

since, as most __dplyr__ functions, `summarize` always returns a data frame.

This might be problematic if we want to use this result with functions that require a numeric value. Here we show a useful trick for accessing values stored in data when using pipes: when a data object is piped that object and its columns can be accessed using the `pull` function. To understand what we mean take a look at this line of code:

```{r}
us_murder_rate %>% pull(rate)
```

This returns the value in the `rate` column of `us_murder_rate` making it equivalent to `us_murder_rate$rate`.

To get a number from the original data table with one line of code we can type:

```{r}
us_murder_rate <- murders %>% 
  summarize(rate = sum(total)/sum(population) * 100000) %>% 
  pull(rate)

us_murder_rate
```

which is now a numeric:

```{r}
class(us_murder_rate)
```

### 5.7.3 Group then summarize with `group_by`

A common operation in data exploration is to first split data into groups and then compute summaries for each group. For example, we may want to compute the average and standard deviation for men's and women's heights separately. The `group_by` function helps us do this.

If we type this:

```{r}
heights %>% group_by(sex)
```

The result does not look very different from `heights`, except we see `Groups: sex [2]` when we print the object. Although not immediately obvious from its appearance, this is now a special data frame called a _grouped data frame_ and __dplyr__ functions, in particular `summarize`, will behave differently when acting on this object. Conceptually, you can think of this table as many tables, with the same columns but not necessarily the same number of rows, stacked together in one object. When we summarize the data after grouping, this is what happens:

```{r}
heights %>% 
  group_by(sex) %>% 
  summarize(average = mean(height), standard_deviation = sd(height))
```

The `summarize` function applies the summarization to each group separately.

For another example, let’s compute the median murder rate in the four regions of the country:

```{r}
murders %>% 
  group_by(region) %>% 
  summarize(median_rate = median(rate))
```

## 5.8 Sorting data frames

When examining a dataset, it is often convenient to sort the table by the different columns. We know about the `order` and `sort` function, but for ordering entire tables, the __dplyr__ function `arrange` is useful. For example, here we order the states by population size:

```{r}
murders %>% 
  arrange(population) %>% 
  head()
```

```{r}
murders %>% 
  arrange(-population) %>% 
  head()
```

With `arrange` we get to decide which column to sort by. To see the states by murder rate, from smallest to largest, we arrange by `rate` instead:

```{r}
murders %>% 
  arrange(rate) %>% 
  head()
```

Note that the default behavior is to order in ascending order. In __dplyr__, the function `desc` transforms a vector so that it is in descending order. To sort the table in descending order, we can type:

```{r}
murders %>% 
  arrange(desc(rate)) %>% 
  head()
```

```{r}
murders %>% 
  arrange(-rate) %>% 
  head()
```

### 5.8.1 Nested sorting

If we are ordering by a column with ties, we can use a second column to break the tie. Similarly, a third column can be used to break ties between first and second and so on. Here we order by `region`, then within region we order by murder rate:

```{r}
murders %>% 
  arrange(region, rate) %>% 
  head()
```

### 5.8.2 The top $n$

In the code above, we have used the function `head` to avoid having the page fill up with the entire dataset. If we want to see a larger proportion, we can use the `top_n` function. This function takes a data frame as its first argument, the number of rows to show in the second, and the variable to filter by in the third. Here is an example of how to see the top 10 rows:

```{r}
murders %>% top_n(10, rate)
```

Note that rows are not sorted by `rate`, only filtered. If want to sort, we need to use `arrange`. Note that if the third argument is left blank, `top_n` filters by the last column.

Another example:

```{r}
murders %>% 
  top_n(3, population) %>% 
  arrange(-population)
```

## 5.9 Exercises

For these exercises, we will be using the data from the survey collected by the United States National Center for Health Statistics (NCHS). This center has conducted a series of health and nutrition surveys since the 1960's. Starting in 1999, aboutg 5,000 individuals of all ages have been interviewed every year and they complete the health examination component of the survey. Part of the data is made available via the __NHANES__ package. Once you install the __NHANES__ package, you can load the data like this:

```{r}
library(NHANES)
data(NHANES)
```

The __NHANES__ data has many missing values. Remember that the main summarization function in R will return `NA` if any of the entries of the input vector is an `NA`. Here is an example:

```{r}
library(dslabs)
data(na_example)
mean(na_example)
```

```{r}
sd(na_example)
```
To ignore the NAs we can use the na.rm argument:
```{r}
mean(na_example, na.rm = TRUE)
sd(na_example, na.rm = TRUE)
```
Let's now explore the NHANES data.

1. We will provide some basic facts about blood pressure. First let's select a group to set the standard. We will use 20-29 year old females. `AgeDecade` is a categorial variable with these ages. Note that the category is coded like " 20-29", with a space in front! What is the average and standard deviation of systolic blood pressure as saved in the `BPSysAve` variable? Save it to a variable called `ref`.

Hint: Use `filter` and `summarize` and use the `na.rm = TRUE` argument when computing the average and standard deviation. You can also filter the NA values using `filter`.

```{r}
#Filter NA values using mean() and sd() argument `na.rm = TRUE`
ref <- NHANES %>% 
  filter(AgeDecade == " 20-29") %>% 
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
ref
```
_Note that `na.rm` is NOT a function but an argument._


```{r}
#Filter NA values using filter and is.na
ref_isna <- NHANES %>% 
  filter(AgeDecade == " 20-29" & is.na(BPSysAve) == FALSE) %>% 
  summarize(average = mean(BPSysAve), standard_deviation = sd(BPSysAve))
ref_isna
```


2. Using a pipe, assign the average to a numeric variable `ref_avg`. 

Hint: Use the code similar to above and then `pull`.

다음과 같이 하면 안된다. Pipe 쓰라했음.

```{r}
ref_avg <- pull(ref, average)
ref_avg
```
문제가 요구하는대로 pipe를 이용해 풀면,
```{r}
ref_avg <- ref %>% pull(average)
ref_avg
class(ref_avg)
```

3. Now report the min and max values for the same group.

```{r}
ref_min_max <- NHANES %>% filter(AgeDecade == " 20-29") %>% summarize(minimum = min(BPSysAve, na.rm = TRUE), maximum = max(BPSysAve, na.rm = TRUE))
ref_min_max

ref_min <- ref_min_max %>% pull(minimum)
ref_max <- ref_min_max %>% pull(maximum)
```

4. Compute the average and standard deviation for females, but for each age group separately rather than a selected decade as in question 1. Note that the age groups are defined by `AgeDecade`. Hint: rather than filtering by age and gender, filter by `Gender` and then use `group_by`.

```{r}
NHANES %>% 
  filter(Gender == "female" & !is.na(AgeDecade)) %>% 
  group_by(AgeDecade) %>% 
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

What if we contain AgeDecade values of NA?

```{r}
NHANES %>% 
  filter(Gender == "female") %>% 
  group_by(AgeDecade) %>% 
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

5. Repeat exercise 4 for males.

```{r}
NHANES %>% filter(Gender == "male") %>% group_by(AgeDecade) %>% summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

```{r}
NHANES %>% filter(Gender == "male" & !is.na(AgeDecade)) %>% group_by(AgeDecade) %>% summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

6. We can actually combine both summaries for exercises 4 and 5 into one line of code. This is because `group_by` permits us to group by more than one variable. Obtain one big summary table using `group_by(AgeDecade, Gender)`.

```{r}
NHANES %>% filter(!is.na(AgeDecade) & !is.na(Gender)) %>% group_by(AgeDecade, Gender) %>% summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE)) %>% arrange(Gender)
```

7. For males between the ages of 40-49, compare systolic blood pressure across race as reported in the `Race1` variable. Order the resulting table from loweset to highest average systolic blood pressure.

```{r}
head(NHANES$Race1)
```

`Race1` is a factor that consists of 5 levels, namely Black, Hispanic, Mexican, White, and Other.
I'm gonna filter by `Gender` "male" and `AgeDecade` " 40-49", and use `group_by(Race1)` to establish a grouped data frame via race.

```{r}
NHANES %>% filter(Gender == "male" & AgeDecade == " 40-49") %>% group_by(Race1) %>% summarize(average_systolic_BP = mean(BPSysAve, na.rm = TRUE)) %>% arrange(average_systolic_BP)
```
## 5.10 Tibbles

Tidy data must be stored in data frames. We introduced the data frame in Section 3.5 and have been using the `murders` data frame throughout the book:

```{r}
data(murders)
class(murders)
```

In Section 5.7.3 we introduced the `group_by` function, which permits stratifying data before computing summary statistics. But where is the group information stored in the data frame?

```{r}
murders %>% group_by(region) %>% head()
```

Notice that there are no columns with this information. But, if you look closely at the output above, you see the line `A tibble:  6 x 5`. We can learn the class of the returned object using:

```{r}
murders %>% group_by(region) %>% class()
```

The `tbl`, pronounced tibble, is a special kind of data frame. The functions `group_by` and `summarize` always return this type of data frame. The `group_by` function returns a special kind of `tbl`, the `grouped_df`. We will say more about these later. For consistency, the __dplyr__ manipulation verbs (`select`, `filter`, `mutate`, and `arrange`) preserve the class of the input: if they receive a regular data frame they return a regular data frame, while if they receive a tibble they return a tibble. But tibbles are the preferred format in the tidyverse and as a result tidyverse functions that produce a data frame from scratch return a tibble. For example, in Chapter 6 we will see that tidyverse functions used to import data create tibbles.

> Tidyverse는 tibble을 좋아한다. Input이 data frame 이더라도 output 항상 tibble임을 염두에 둬야한다.

Tibbles are very similar to data frames. In fact, you can think of them as a modern version of data frames. Nonetheless there are three important differences which we describe in the next.

### 5.10.1 Tibbles display better

The print method for tibbles is more readable than that of a data frame. To see this, compare the outputs of typing `murders` and the output of murders if we convert it to a tibble. We can do this using `as_tibble(murders)`. If using RStudio, output for a tibble adjusts to your window size. To see this, change the width of your R console and notice how more/less columns are shown.

```{r, eval= FALSE, echo = FALSE}
murders
```

```
                  state abb        region population total
1               Alabama  AL         South    4779736   135
2                Alaska  AK          West     710231    19
3               Arizona  AZ          West    6392017   232
4              Arkansas  AR         South    2915918    93
5            California  CA          West   37253956  1257
6              Colorado  CO          West    5029196    65
7           Connecticut  CT     Northeast    3574097    97
8              Delaware  DE         South     897934    38
9  District of Columbia  DC         South     601723    99
10              Florida  FL         South   19687653   669
11              Georgia  GA         South    9920000   376
12               Hawaii  HI          West    1360301     7
13                Idaho  ID          West    1567582    12
14             Illinois  IL North Central   12830632   364
15              Indiana  IN North Central    6483802   142
16                 Iowa  IA North Central    3046355    21
17               Kansas  KS North Central    2853118    63
18             Kentucky  KY         South    4339367   116
...
```
```{r, eval = FALSE, echo = FALSE}
as_tibble(murders)
```
```
# A tibble: 51 x 5
   state                abb   region    population total
   <chr>                <chr> <fct>          <dbl> <dbl>
 1 Alabama              AL    South        4779736   135
 2 Alaska               AK    West          710231    19
 3 Arizona              AZ    West         6392017   232
 4 Arkansas             AR    South        2915918    93
 5 California           CA    West        37253956  1257
 6 Colorado             CO    West         5029196    65
 7 Connecticut          CT    Northeast    3574097    97
 8 Delaware             DE    South         897934    38
 9 District of Columbia DC    South         601723    99
10 Florida              FL    South       19687653   669
# … with 41 more rows
```

### 5.10.2 Subsets of tibbles are tibbles

If you subset the columns of a data frame, you may get back an object that is not a data frame, such as a vector or scalar. For example:

```{r}
class(murders[,4])
```

is not a data frame. With tibbles this does not happen:

```{r}
class(as_tibble(murders)[,4])
```

This is userful in the tidyverse since functions require data frames as input.

With tibbles, if you want to access the vector that defines a column, and not get back a data frame, you need to use the accessor `$`:

```{r}
class(as_tibble(murders)$population)
```

A related feature is that tibbles will give you a warning if you try to access a column that does not exist. If we accidentally write `Population` instead of `population` this:
```{r}
murders$Population
```
returns a `NULL` with no warning, which can make it harder to debug. In contrast, if we try this with a tibble we get an informative warning:

```{r}
as_tibble(murders)$Population
```
In the console:
```
NULL
Warning message:
Unknown or uninitialised column: 'Population'. 
```

### 5.10.3 Tibbles can have complex entries

While data frame columns need to be vectors of numbers, strings or logical values, tibbles can have more complex objects, such as lists or functions. Also, we can create tibbles with functions:

```{r, eval = FALSE, echo = FALSE}
tibble(id = c(1, 2, 3), func = c(mean, median, sd))
```

```
# A tibble: 3 x 2
     id func  
  <dbl> <list>
1     1 <fn>  
2     2 <fn>  
3     3 <fn>  
```

### 5.10.4 Tibbles can be grouped

The function `group_by` returns a special kind of tibble: a grouped tibble. This class stores information that lets you know which rows are in which groups. The tidyverse functions, in particular the `summarize` function, are aware of the group information.

### 5.10.5 Create a tibble using `data_frame` instead of `data.frame`
It is sometimes useful for us to create our own data frames. To create a data frame in the tibble format, you can do this by using the `data_frame` function.
> data.frame은 base function이고, data_frame은 dplyr library에 포함된 function이다.

```{r}
grades <- data_frame(names = c("John", "Juan", "Jean", "Yao"), 
                     exam_1 = c(95, 80, 90, 85), 
                     exam_2 = c(90, 85, 85, 90))
```
ㄷㄷㄷ... 뭐지 이런 warning이 뜨는 이유는...? ㄷㄷ
오래된 문법으로 취급되나보다. 더 이상 사용되지 않는다고 한다. 이제 `tibble()`로 통일하라는 뜻인것 같다.
```{r}
grades
```
tibble로 해보자.
```{r}
grades2 <- tibble(subjects = c("korean", "math", "english"),
                  kim = runif(3)*100,
                  oh = runif(3)*100,
                  park = runif(3)*100)
grades2
```


```{r}
grades2 %>% group_by(subjects) %>% summarize(average_subject = mean(subjects))
```
지금 이 접근방식이 안되는 이유는 이게 tidy format이 아니기 때문이다.
Tidy format의 tibble을 하나 시험삼아 만들어보자.

```{r}
grades_tidy <- tibble(name = factor(c("kim", "kim", "kim", "oh", "oh", "oh", "park", "park", "park")),
       subject = factor(c("korean","math","english","korean","math","english","korean","math", "english")),
       grade = runif(9)*100)
grades_tidy
```

학생과 과목별 평균을 구해보자.
```{r}
grades_tidy %>% group_by(name) %>% summarize(average_by_student = mean(grade))
grades_tidy %>% group_by(subject) %>%  summarize(average_by_subject = mean(grade))
```
영어에서 2등한 학생을 찾아보자.

```{r}
grades_tidy %>% filter(subject == "english") %>% mutate(rank = rank(grade)) %>% filter(rank == 2)
```

Note that base R (without packages loaded) has a function with a very similar name, `data.frame`, that can be used to create a regular data frame rather than a tibble. One other important difference is that by default `data.frame` **coerces** characters into factors without providing a warning or message:

```{r}
grades <- data.frame(names = c("John", "Juan", "Jean", "Yao"), 
                     exam_1 = c(95, 80, 90, 85), 
                     exam_2 = c(90, 85, 85, 90))
class(grades$names)
```

To avoid this, we use the rather cumbersome argument `stringsAsFactors`:
```{r}
grades <- data.frame(names = c("John", "Juan", "Jean", "Yao"), 
                     exam_1 = c(95, 80, 90, 85), 
                     exam_2 = c(90, 85, 85, 90),
                     stringsAsFactors = FALSE)
class(grades$names)
```

To convert a regular data frame to a tibble, you can use the `as_tibble` function.
```{r}
as_tibble(grades) %>% class()
```

## 5.11 The dot operator

One of the advantages of using the pipe `%>%` is that we do not have to keep naming new objects as we manipulate the data frame. As quick reminder, if we want to compute the median murder rate for states in the southern states, instead of typing:

```{r}
tab_1 <- filter(murders, region == "South")
tab_2 <- mutate(tab_1, rate = total / population * 10^5)
rates <- tab_2$rate
median(rates)
```

We can avoid defining any new intermediate objects by instead typing:

```{r}
filter(murders, region == "South") %>% 
  mutate(rate = total / population * 10^5) %>% 
  summarize(median = median(rate)) %>% 
  pull(median)
```

We can do this because each of these functions takes a data frame as the first argument. But what if we want to access a component of the data frame. For example, what if the `pull` function was not available and we wanted to access `tab_2$rate`? What data frame name would we use? The answer is the dot operator.

For example to access the rate vector without the `pull` function we could use
```{r}
rates <- filter(murders, region == "South") %>% 
  mutate(rate = total / population * 100000) %>% 
  .$rate
median(rates)
```

In the next section, we will see other instances in which using the `.` is useful.

